---
title: "Project 1 Reproducible Research"
output: 
  html_document: 
    keep_md: yes
---


## PART 1 Loading and preprocessing the data

```{r Code to Clean and Construct Data Frames, echo=TRUE}
Data<-read.csv("activity.csv")
DataClean<-na.omit(Data)
library(plyr)
summary(DataClean)
```
Note that all NAs were removed but we still have zero min quantity

## Part II What is mean total number of steps taken per day?

Histogram of the total number of steps the individuals in our data set 
```{r Histogram of Steps, echo=TRUE}
SumData<-aggregate(DataClean["steps"],list(date = DataClean$date),sum)
hist(SumData$steps, breaks=50, main="Steps by Individual")
summary(SumData)
```
Note the mean and median from the summary taken on the data set

## Part III What is the average daily activity pattern?

Plot of the Daily Steps Taken (in 5 Minute intervals) 
```{r Daily Acitivity Pattern, echo=TRUE}
MeanData<-aggregate(DataClean["steps"],list(Interval = DataClean$interval),mean)
plot(MeanData$Interval, MeanData$steps, type ="l", main= "Mean Steps by Interval Period")
summary(MeanData)
```
See summary above for max interval of 2355

## Part IV Imputing missing values

How many values of NA do I have in my original Dataset?
```{r Data Set NA evaluation, echo=TRUE}
StepsNA<-sum(is.na(Data$steps))
DataNA<-sum(is.na(Data$date))
IntervalNA<-sum(is.na(Data$interval))
```

There are `r StepsNA` rows containing an NA value, all of which are the the Steps column

```{r Find and Replace NA, echo=TRUE}
## First append DataMean to Data based on matching interval
Data$match<-MeanData$steps[match(Data$interval,MeanData$Interval)]
## Then conditionally index the Data Frame and replace all NA in step column 1 with match column 4
Index<-is.na(Data$steps)
Data$steps[Index]<-Data$match[Index]
##Verify NA's have been replaced by running sum(is.na(Data$Steps))
SumDataNEW<-aggregate(Data["steps"],list(date = Data$date),sum)
hist(SumDataNEW$steps, breaks=50, main="Steps by Individual")
summary(SumDataNEW)
summary(SumData)
```

When comparing the mean and median of the old data set and the new data set, per the summary of each, the totals work out to be the same.
The impact is in the quantile data and in the total number of records

## Part V Are there differences in activity patterns between weekdays and weekends?

```{r Daily Acitivity Pattern by weekday or weekend, echo=TRUE}
## Make sure your date column changes from a 'factor' class to a 'date' class
DataClean$date<-as.Date(DataClean$date)
## Add new factor column using weekdays() function
DataClean$Weekdata<-weekdays(DataClean$date)
## Add Another factor column to state whether the day is a weekend or weekday
DataClean$Day_or_End<- ifelse(DataClean$Weekdata == "Saturday" | DataClean$Weekdata == "Sunday", "Weekend", "Weekday")
##Final Aggregate data to plot
FinalAGG<-aggregate(steps ~ interval + Day_or_End, DataClean, sum)
attach(FinalAGG)
par(mfrow=c(2,1))
plot(subset(FinalAGG, Day_or_End == "Weekend", select=c("interval","steps")), type="l", main="Weekend steps by interval")
plot(subset(FinalAGG, Day_or_End == "Weekday", select=c("interval","steps")), type="l", main="Weekday steps by interval")
```
knit2html()
